---
title: "Prior Exploration For Simulations"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
```{r}
library(ggplot2)
library(ggsci)
library(ggpubr)
library(rethinking)
library(dplyr)
library(gridGraphics)
```


```{r}
calc_direct_V<-function(phy, sigma2_y, a){
  ts<-ts_fxn(phy)
  ta<-ts[[1]]
  tij<-ts[[2]] #Same as Cophenetic Distance matrix
  Vt<-sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) * exp(-a * tij)) #ta - time from root to tips, tij  - total time separating spcies
  return(Vt)
}
```

```{r}
calc_adaptive_V<-function(a, sigma2_y, ta,  tij,  tja,  T_term,  beta,  sigma2_x){
  N<-dim(ta)[1];
  Z<-length(beta);
  ts<-ts_fxn(phy)
  ta<-ts[[1]]
  tij<-ts[[2]] #Same as Cophenetic Distance matrix
  T_term<-ts[[3]]
  tja<-ts[[4]]
  ones<-rep(1,Z)
  ti<-matrix(T_term,length(T_term),N);
  if(Z==1){var_opt<-beta^2*sigma2_x[1,1]}
  else(var_opt<-as.numeric(matrix(beta,nrow=1,ncol=Z)^2 %*% sigma2_x %*% ones))
  term0 = ((var_opt + sigma2_y) / (2 * a)) * (1 - exp( -2 * a * ta)) * exp(-a * tij)
  term1 = (1 - exp(-a * ti)) / (a * ti)
  term2 = exp(-a * tja) * (1 - exp(-a * ti)) / (a * ti)
  Vt<-term0 + var_opt * (ta * term1 * t(term1) - ((1 - exp(-a * ta)) / a) * (term2 + t(term2))) 
  return(Vt)
}
```

```{r}
calc_adaptive_cov<-function(a,sigma2_y,beta,x){
  ti<-1
  Z<-length(beta)
  if(Z==1){var_opt<-beta^2
  }else{var_opt<-t(beta)%*%matrix(1,Z,Z)%*%c(rep(1,Z))}
  term0 = ((var_opt + sigma2_y) / (2 * a)) * (1 - exp( -2 * a * (1-x/2))) * exp(-a * x)
  term1 = (1 - exp(-a * ti)) / (a * ti)
  term2 = exp(-a * x) * (1 - exp(-a * ti)) / (a * ti)
  Vt<-term0 + var_opt * ((1-x/2) * term1 * term1 - ((1 - exp(-a * (1-x/2))) / a) * (term2 + term2)) 
  return(Vt)
}
```


```{r}
ts_fxn<-function(phy){ #Calculate t
  n<-length(phy$tip.label)
  mrca1 <- ape::mrca(phy) #Node numbers for MRCA of tips
  times <- ape::node.depth.edgelength(phy) #Time from root to tips of each node, starting with the tips
  ta <- matrix(times[mrca1], nrow=n, dimnames = list(phy$tip.label, phy$tip.label)) #Matrix with time from root to MRCA of pairs of tips - pulls out values of times that correspond with node numbers - integers
  T.term <- times[1:n] #Times from root for tips
  tia <- times[1:n] - ta #Times from root to tips - times from root to MRCA = times from MRCA to tips
  tja <- t(tia) #Transpose of the times from MRCA to tips matrix
  #tij <- tja + tia #Sum of matrix and its transpose - total time separating species
  tij<-cophenetic(phy)
  #return(list(ta,tia,tja,tij,T.term))
  return(list(ta,tij,T.term,tja))
}
```
  

*******************

Basic direct effect and adaptation model plots
#########################################################
Milestones 1,2: Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}

hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
names(hl.sims)<-"prior.hl.sims"


hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"


hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+
  
  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=5)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+
  
  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```


Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,1) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```


Adaptation model
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,1) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(calc_adaptive_cov(a.sims[i,],sigma2_y.sims[i,],beta,x) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(calc_adaptive_cov(post$a[i],post$sigma2_y[i],post$beta[i],x) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

#multiple traits


par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```


Slope and intercept - simple one level no regimes model
Prior vs. Posterior Plot
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

#Adaptive Model
alpha.sims<-rnorm(100,2,0.2)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-data.frame(post$alpha)
names(alpha.post)<-"post.alpha"

beta.post<-data.frame(post$beta)
names(beta.post)<-"post.beta"

mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu <- sapply(x.seq , mu.link )
mu.mean <-lapply( mu , mean )
mu.mean<-data.frame(as.numeric(mu.mean))
names(mu.mean)<-"mu.mean"

mu.CI <- lapply( mu , PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),x.seq)
names(mu.CI)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,X=X)
names(df)<-c("Y","X")

df2<-data.frame(x.seq,mu.mean)

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+
  geom_abline(intercept=alpha,slope=beta,alpha=0.5,linetype=2)+
  geom_line(data=df2,aes(x=x.seq,y=mu.mean),linetype=1)+
  geom_ribbon(data=mu.CI,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+
  
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Trait 1") + xlab("Trait 2")+
  scale_color_npg()

slope.plot
```


Change figure number depending on model/parameter combination
```{r}
#text.box<-paste("Fig. S12: Prior vs. posterior for A) Half-life, B) vy, and C) Predicted mean intercept and slope and 89% compatibility interval for the means with priors in light grey and posteior in dark grey colors. For this simulation parameter values were set to: hl=0.75, vy=0.1, alpha=2, beta=0.25; Priors: hl~lognormal(log(0.25),0.75); vy~exponential(5); alpha~normal(0,1); beta~normal(0,1). Dotted lines in A) and B) are true values of the parameter")
#text.box <- ggparagraph(text.box, face = "plain", size = 11)

fig<-ggarrange(hl.plot, vy.plot, "",slope.plot, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS13.pdf", plot = fig, width=7, height=7 )

#ggsave("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS10.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS13C.pdf",   # The directory you want to save the file in
    width = 3.57, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches


#pdf(file = "/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS10C.pdf",   # The directory you want to save the file in
#    width = 3.57, # The width of the plot in inches
#    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
  

```


####
Multiple traits
```{r}
library(ggsci)
library(rethinking)

alpha.sims<-rnorm(100,2,0.2)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-data.frame(post$alpha)
names(alpha.post)<-"post.alpha"

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

beta.post.2<-data.frame(post$beta[,2])
names(beta.post.2)<-"post.beta.2"

mu.link.1<-function(x.seq){alpha.post+x.seq*beta.post.1}
mu.link.2<-function(x.seq){alpha.post+x.seq*beta.post.2}
x.seq <- seq(from=min(Xs), to=max(Xs) , length.out=100)
mu.1 <- sapply(x.seq , mu.link.1 )
mu.2 <- sapply(x.seq , mu.link.2 )
mu.mean.1 <-lapply( mu.1 , mean )
mu.mean.2 <-lapply( mu.2 , mean )
mu.mean.1<-data.frame(as.numeric(mu.mean.1))
mu.mean.2<-data.frame(as.numeric(mu.mean.2))
names(mu.mean.1)<-"mu.mean.1"
names(mu.mean.2)<-"mu.mean.2"

mu.CI.1 <- lapply( mu.1 , PI , prob=0.89 )
mu.CI.2 <- lapply( mu.2 , PI , prob=0.89 )
mu.CI.1<-data.frame(t(data.frame(mu.CI.1)),x.seq)
mu.CI.2<-data.frame(t(data.frame(mu.CI.2)),x.seq)

names(mu.CI.1)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.2)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs)
df2<-data.frame(x.seq,mu.mean.1)
df3<-data.frame(x.seq,mu.mean.2)

#Combined plots
#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
#slope.plot.1<-ggplot()+  
#  geom_point(data=df,aes(y=Y,x=X.1))+
#  geom_point(data=df,aes(y=Y,x=X.2),alpha=0.25)+
#  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
#  geom_line(data=df2,aes(x=x.seq,y=mu.mean.1),linetype=1)+
#  geom_ribbon(data=mu.CI.1,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
#  geom_line(data=df3,aes(x=x.seq,y=mu.mean.2),linetype=1,alpha=0.5)+
#  geom_ribbon(data=mu.CI.2,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+

#  theme_bw()+
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
#  ylab("Trait 1") + xlab("Trait 2")+
#  scale_color_npg()

#slope.plot.1


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.1))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=alpha,slope=beta[1],alpha=0.5,linetype=2)+
  geom_line(data=df2,aes(x=x.seq,y=mu.mean.1),linetype=1)+
  geom_ribbon(data=mu.CI.1,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+
  
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait 1")+
  scale_color_npg()

slope.plot.1

slope.plot.2<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.2))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=alpha,slope=beta[2],alpha=0.5,linetype=2)+
  geom_line(data=df3,aes(x=x.seq,y=mu.mean.2),linetype=1)+
  geom_ribbon(data=mu.CI.2,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait 2")+
  scale_color_npg()

slope.plot.2

```



```{r}
fig<-ggarrange(hl.plot, vy.plot, "",slope.plot.1, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS14.pdf", plot = fig, width=7, height=7 )

fig<-ggarrange(slope.plot.2, ncol=1,nrow=1, labels = c("E)"))

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS14E.pdf", plot = fig, width=3.5, height=3.3)


pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS14C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
```


#################################################################################
Milestones 3: Combination direct effect and adaptation models
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"


hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=5)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```


Adaptation model
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,1) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(calc_adaptive_cov(a.sims[i,],sigma2_y.sims[i,],beta[2],x) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(calc_adaptive_cov(post$a[i],post$sigma2_y[i],post$beta[i,2],x) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```

Slope plots for direct and adaptive models
```{r}
library(ggsci)
library(rethinking)

alpha.sims<-rnorm(100,2,0.2)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-data.frame(post$alpha)
names(alpha.post)<-"post.alpha"

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

beta.post.2<-data.frame(post$beta[,2])
names(beta.post.2)<-"post.beta.2"

mu.link.1<-function(x.seq){alpha.post+x.seq*beta.post.1}
mu.link.2<-function(x.seq){alpha.post+x.seq*beta.post.2}
x.seq <- seq(from=min(Xs), to=max(Xs) , length.out=100)
mu.1 <- sapply(x.seq , mu.link.1 )
mu.2 <- sapply(x.seq , mu.link.2 )
mu.mean.1 <-lapply( mu.1 , mean )
mu.mean.2 <-lapply( mu.2 , mean )
mu.mean.1<-data.frame(as.numeric(mu.mean.1))
mu.mean.2<-data.frame(as.numeric(mu.mean.2))
names(mu.mean.1)<-"mu.mean.1"
names(mu.mean.2)<-"mu.mean.2"

mu.CI.1 <- lapply( mu.1 , PI , prob=0.89 )
mu.CI.2 <- lapply( mu.2 , PI , prob=0.89 )
mu.CI.1<-data.frame(t(data.frame(mu.CI.1)),x.seq)
mu.CI.2<-data.frame(t(data.frame(mu.CI.2)),x.seq)

names(mu.CI.1)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.2)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs)
df2<-data.frame(x.seq,mu.mean.1)
df3<-data.frame(x.seq,mu.mean.2)

#Combined plots
#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
#slope.plot.1<-ggplot()+  
#  geom_point(data=df,aes(y=Y,x=X.1))+
#  geom_point(data=df,aes(y=Y,x=X.2),alpha=0.25)+
#  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
#  geom_line(data=df2,aes(x=x.seq,y=mu.mean.1),linetype=1)+
#  geom_ribbon(data=mu.CI.1,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
#  geom_line(data=df3,aes(x=x.seq,y=mu.mean.2),linetype=1,alpha=0.5)+
#  geom_ribbon(data=mu.CI.2,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+

#  theme_bw()+
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
#  ylab("Trait 1") + xlab("Trait 2")+
#  scale_color_npg()

#slope.plot.1


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.1))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=alpha,slope=beta[1],alpha=0.5,linetype=2)+
  geom_line(data=df2,aes(x=x.seq,y=mu.mean.1),linetype=1)+
  geom_ribbon(data=mu.CI.1,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1

slope.plot.2<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.2))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=alpha,slope=beta[2],alpha=0.5,linetype=2)+
  geom_line(data=df3,aes(x=x.seq,y=mu.mean.2),linetype=1)+
  geom_ribbon(data=mu.CI.2,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait")+
  scale_color_npg()

slope.plot.2

```


Change figure number depending on model/parameter combination
```{r}
#text.box<-paste("Fig. S24: Prior vs. posterior for A) Half-life, B) vy, and C,D) Posterior prediction means (black line) and 89% compatibility intervals (light grey region) for direct effect and adaptive models, with priors in light grey. Species values are shown in the dark circles. For this simulation parameter values were set to: hl=0.1, vy=0.01, alpha=2, beta=(0.35,0.25); Priors: hl~lognormal(log(0.25),0.75); vy~exponential(5); alpha~normal(2,0.2); beta~normal(0,0.25). Dotted lines in A) and B) are true values of the parameter")
#text.box <- ggparagraph(text.box, face = "plain", size = 11)

fig<-ggarrange(hl.plot, vy.plot, "",slope.plot.1, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS20.pdf", plot = fig, width=7, height=7 )

fig<-ggarrange(slope.plot.2, ncol=1,nrow=1, labels = c("E)"))

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS20E.pdf", plot = fig, width=3.5, height=3.3)


pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS20C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()

#fig<-ggarrange(slope.plot.2, ncol=1,nrow=1, labels = c("E)"))

#ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS14E.pdf", plot = fig, width=3.5, height=3.3)


#fig<-ggarrange(hl.plot, vy.plot, slope.plot.1, slope.plot.2, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

#fig<-ggarrange(fig, text.box, ncol=1,nrow=2)

#ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figSX.pdf", plot = fig, width=7, height=7 )

#pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figSXD.pdf",   # The directory you want to save the file in
#    width = 3.85, # The width of the plot in inches
#    height = 3.65) # The height of the plot in inches
#covariance.plot

dev.off()


```



#############################################################################################################################Milestones 4: Regime Model
Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"


hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"


hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
      theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=30)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
      theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```


Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.3) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```


Prior vs. Posterior Plot
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

beta.sims<-rnorm(n=100,0,1)
beta.sims<-data.frame(beta.sims,Prior="Prior")
optima.post<-data.frame(post$optima)
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,optima)
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=beta.sims,aes(y=beta.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),size=35,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
#      theme(
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank())+

  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```

Change figure number depending on model/parameter combination
```{r}
#text.box<-paste("Fig. S19: Prior vs. posterior for A) Half-life, B) vy, and C) Mean prediceted optima shown in solid lines and 89% compatibility interval for the means in shaded area, with simulated species values in dark great and priors in light grey circles. Dotted lines overlaying estimated optima are true values of the parameter. For this simulation parameter values were set to: hl=0.75, vy=0.01, optima=(0.5,0.25); Priors: hl~lognormal(log(0.25),0.75); vy~exponential(20); beta~normal(0,1).")
#text.box <- ggparagraph(text.box, face = "plain", size = 11)

fig<-ggarrange(hl.plot, vy.plot, covariance.plot, optima.plot, ncol=2,nrow=2, labels = c("A)","B)","C)"),common.legend = TRUE,legend="top")
#ggsave("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS17.pdf", plot = fig, width=7, height=7 )
ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS27.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS27C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()


```


###############################################################################################
Milestone 5: Regime + Direct Effect Plots

Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"


hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"


hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```



Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.5) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```


Slope plots for direct model without measurement error
```{r}
library(ggsci)
library(rethinking)

alpha.sims<-rnorm(100,1.5,0.25)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

mu.link.11<-function(x.seq){alpha.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){alpha.post[,2]+x.seq*beta.post.1}
x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"

mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)


names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=beta[1],slope=beta[3],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[3],alpha=0.5,linetype=2)+
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1


```

Slope plots for direct model with measurement error
```{r}
library(ggsci)
library(rethinking)
X<-X_with_error
Y<-Y_with_error

alpha.sims<-rnorm(100,1.5,0.25)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

mu.link.11<-function(x.seq){alpha.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){alpha.post[,2]+x.seq*beta.post.1}
x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"

mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)


names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=beta[1],slope=beta[3],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[3],alpha=0.5,linetype=2)+
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1


```



Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, "",slope.plot.1, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS32.pdf", plot = fig, width=7, height=7 )


pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS32C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
```


###############################################################################################
Milestone 6: Regime + Adaptive Plots
Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"


hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"


hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
      theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```


```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
      theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```

Adaptation model
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.4) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(calc_adaptive_cov(a.sims[i,],sigma2_y.sims[i,],beta[3],x) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(calc_adaptive_cov(post$a[i],post$sigma2_y[i],post$beta[i],x) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

#multiple traits


par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```


Prior vs. Posterior Plot
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

beta.sims<-rnorm(n=100,2,1)
beta.sims<-data.frame(beta.sims,Prior="Prior")
optima.post<-data.frame(post$optima[,1:2])
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,"optima"=beta[1:2])
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=beta.sims,aes(y=beta.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),size=35,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
      theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```

Slope plots for direct and adaptive models without error
```{r}
library(ggsci)
library(rethinking)

alpha.sims<-rnorm(100,1.5,0.5)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"


mu.link.11<-function(x.seq){alpha.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){alpha.post[,2]+x.seq*beta.post.1}
x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"

mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)


names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
#df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip) #For error models
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=beta[1],slope=beta[3],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[3],alpha=0.5,linetype=2)+
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait")+
  scale_color_npg()

slope.plot.1

```



Slope plots for direct and adaptive models with error
```{r}
library(ggsci)
library(rethinking)
X<-X_with_error #For error models
Y<-Y_with_error #For error models

alpha.sims<-rnorm(100,1.5,0.5)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"


mu.link.11<-function(x.seq){alpha.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){alpha.post[,2]+x.seq*beta.post.1}
x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"

mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)


names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
#df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df<-data.frame(Y=Y,X=X,Regimes=regimes_tip) #For error models
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=beta[1],slope=beta[3],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[3],alpha=0.5,linetype=2)+
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait")+
  scale_color_npg()

slope.plot.1

```



Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, "",slope.plot.1, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS38.pdf", plot = fig, width=7, height=7 )


pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS38C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
```


###############################################################################################
Milestone 7: Regime + Direct + Adaptive Plots
Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"


hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"


hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
          theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
          theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```


Adaptation model
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.4) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(calc_adaptive_cov(a.sims[i,],sigma2_y.sims[i,],beta[3],x) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(calc_adaptive_cov(post$a[i],post$sigma2_y[i],post$beta[i],x) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

#multiple traits


par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```




Prior vs. Posterior Plot
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

beta.sims<-rnorm(n=100,2,1)
beta.sims<-data.frame(beta.sims,Prior="Prior")
optima.post<-data.frame(post$optima[,1:2])
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,"optima"=beta[1:2])
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=beta.sims,aes(y=beta.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),size=35,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```

Slope plots for direct and adaptive models without Measurement Error
```{r}
library(ggsci)
library(rethinking)

alpha.sims<-rnorm(100,1.5,0.25)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

beta.post.2<-data.frame(post$beta[,2])
names(beta.post.2)<-"post.beta.2"

mu.link.11<-function(x.seq){alpha.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){alpha.post[,2]+x.seq*beta.post.1}
mu.link.21<-function(x.seq){alpha.post[,1]+x.seq*beta.post.2}
mu.link.22<-function(x.seq){alpha.post[,2]+x.seq*beta.post.2}
x.seq <- seq(from=min(Xs), to=max(Xs) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.21 <- sapply(x.seq , mu.link.21 )
mu.22 <- sapply(x.seq , mu.link.22 )
mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )
mu.mean.21 <-lapply( mu.21 , mean )
mu.mean.22 <-lapply( mu.22 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"
mu.mean.21<-data.frame(as.numeric(mu.mean.21))
mu.mean.22<-data.frame(as.numeric(mu.mean.22))
names(mu.mean.21)<-"mu.mean.21"
names(mu.mean.22)<-"mu.mean.22"

mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)

mu.CI.21 <- lapply( mu.21 , PI , prob=0.89 )
mu.CI.22 <- lapply( mu.22 , PI , prob=0.89 )
mu.CI.21<-data.frame(t(data.frame(mu.CI.21)),x.seq)
mu.CI.22<-data.frame(t(data.frame(mu.CI.22)),x.seq)

names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.21)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.22)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)
df21<-data.frame(x.seq,mu.mean.21)
df22<-data.frame(x.seq,mu.mean.22)
#Combined plots
#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
#slope.plot.1<-ggplot()+  
#  geom_point(data=df,aes(y=Y,x=X.1))+
#  geom_point(data=df,aes(y=Y,x=X.2),alpha=0.25)+
#  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
#  geom_line(data=df2,aes(x=x.seq,y=mu.mean.1),linetype=1)+
#  geom_ribbon(data=mu.CI.1,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
#  geom_line(data=df3,aes(x=x.seq,y=mu.mean.2),linetype=1,alpha=0.5)+
#  geom_ribbon(data=mu.CI.2,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+

#  theme_bw()+
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
#  ylab("Trait 1") + xlab("Trait 2")+
#  scale_color_npg()

#slope.plot.1


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.1,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_abline(intercept=beta[1],slope=beta[3],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[3],alpha=0.5,linetype=2)+

  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
          theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1

slope.plot.2<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.2,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=beta[1],slope=beta[4],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[4],alpha=0.5,linetype=2)+

  geom_line(data=df21,aes(x=x.seq,y=mu.mean.21),linetype=1)+
  geom_ribbon(data=mu.CI.21,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df22,aes(x=x.seq,y=mu.mean.22),linetype=1)+
  geom_ribbon(data=mu.CI.22,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
          theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait")+
  scale_color_npg()

slope.plot.2

```



Slope plots for direct and adaptive models with Measurement Error
```{r}
library(ggsci)
library(rethinking)

#Y<-Y_with_error #For error models
#X<-X_with_error #For error models


alpha.sims<-rnorm(100,1.5,0.25)
beta.sims<-rnorm(n=100,0,0.25)

alpha.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

beta.post.2<-data.frame(post$beta[,2])
names(beta.post.2)<-"post.beta.2"

mu.link.11<-function(x.seq){alpha.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){alpha.post[,2]+x.seq*beta.post.1}
mu.link.21<-function(x.seq){alpha.post[,1]+x.seq*beta.post.2}
mu.link.22<-function(x.seq){alpha.post[,2]+x.seq*beta.post.2}
x.seq <- seq(from=min(Xs), to=max(Xs) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.21 <- sapply(x.seq , mu.link.21 )
mu.22 <- sapply(x.seq , mu.link.22 )
mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )
mu.mean.21 <-lapply( mu.21 , mean )
mu.mean.22 <-lapply( mu.22 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"
mu.mean.21<-data.frame(as.numeric(mu.mean.21))
mu.mean.22<-data.frame(as.numeric(mu.mean.22))
names(mu.mean.21)<-"mu.mean.21"
names(mu.mean.22)<-"mu.mean.22"



mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)

mu.CI.21 <- lapply( mu.21 , PI , prob=0.89 )
mu.CI.22 <- lapply( mu.22 , PI , prob=0.89 )
mu.CI.21<-data.frame(t(data.frame(mu.CI.21)),x.seq)
mu.CI.22<-data.frame(t(data.frame(mu.CI.22)),x.seq)

names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.21)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.22)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)
df21<-data.frame(x.seq,mu.mean.21)
df22<-data.frame(x.seq,mu.mean.22)
#Combined plots
#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
#slope.plot.1<-ggplot()+  
#  geom_point(data=df,aes(y=Y,x=X.1))+
#  geom_point(data=df,aes(y=Y,x=X.2),alpha=0.25)+
#  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
#  geom_line(data=df2,aes(x=x.seq,y=mu.mean.1),linetype=1)+
#  geom_ribbon(data=mu.CI.1,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
#  geom_line(data=df3,aes(x=x.seq,y=mu.mean.2),linetype=1,alpha=0.5)+
#  geom_ribbon(data=mu.CI.2,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+

#  theme_bw()+
  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
#  ylab("Trait 1") + xlab("Trait 2")+
#  scale_color_npg()

#slope.plot.1


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.1,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_abline(intercept=beta[1],slope=beta[3],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[3],alpha=0.5,linetype=2)+


  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1

slope.plot.2<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X.2,color=Regimes))+
  geom_abline(intercept=alpha.sims,slope=beta.sims,alpha=0.15)+ #Prior
  geom_abline(intercept=beta[1],slope=beta[4],alpha=0.5,linetype=2)+
  geom_abline(intercept=beta[2],slope=beta[4],alpha=0.5,linetype=2)+

  geom_line(data=df21,aes(x=x.seq,y=mu.mean.21),linetype=1)+
  geom_ribbon(data=mu.CI.21,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df22,aes(x=x.seq,y=mu.mean.22),linetype=1)+
  geom_ribbon(data=mu.CI.22,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Adaptive trait")+
  scale_color_npg()

slope.plot.2

```

```{r}
fig1<-ggarrange(hl.plot, vy.plot, ncol=2,nrow=1, labels = c("A)","B)"),common.legend = TRUE,legend="top")

fig2<-ggarrange(slope.plot.1,slope.plot.2, ncol=2,nrow=1, labels = c("C)","D)"),common.legend = TRUE,legend="bottom")

fig<-ggarrange(fig1,fig2, ncol=1,nrow=2)

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS43.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS43E.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
```


#############################################################################################################################
Milestone 8: Multilevel Regime Model with Cenetered priors

Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"

hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```



Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.4) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```



Prior vs. Posterior Plot
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)


optima_bar<-rnorm(n=100,0,1)
#sigma<-abs(rnorm(n=100,0,1))
sigma<-rexp(n=100,rate=5)

optima.sims<-rnorm(n=100,optima_bar,sigma)
optima.sims<-data.frame(optima.sims,Prior="Prior")

optima.post<-data.frame(post$optima)
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,optima)
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=optima.sims,aes(y=optima.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),linewidth=25,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
#            theme(
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank())+

  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```

Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, covariance.plot, optima.plot, ncol=2,nrow=2, labels = c("A)","B)","C)"),common.legend = TRUE,legend="top")
#ggsave("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS17.pdf", plot = fig, width=7, height=7 )
ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS46.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS46C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()


```


#############################################################################################################################
Milestone 8: Multilevel Regime Model with non-cenetered priors

Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"

hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```



Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.4) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```



Prior vs. Posterior Plot
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)


optima_bar<-rlnorm(n=100,0,0.5)
z<-rnorm(n=100,0,1)
sigma<-abs(rnorm(n=100,0,1))
#sigma<-rexp(n=100,rate=5)

optima.sims<-optima_bar + z*sigma;
optima.sims<-data.frame(optima.sims,Prior="Prior")

optima.post<-data.frame(post$optima)
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,optima)
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=optima.sims,aes(y=optima.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),linewidth=25,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
#            theme(
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank())+

  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```

Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, covariance.plot, optima.plot, ncol=2,nrow=2, labels = c("A)","B)","C)"),common.legend = TRUE,legend="top")
#ggsave("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS17.pdf", plot = fig, width=7, height=7 )
ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS47.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS47C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()


```

















#############################################################################################################################
Multilevel Regime Model with non-cenetered priors and SR V/CV

Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.5),sdlog=0.5))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"

hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```

```{r}
var_anc.sims<-rexp(n=1000,rate=100)
var_anc.sims<-data.frame(var_anc.sims)
names(var_anc.sims)<-"prior.var_anc.sims"


var_anc.post<-data.frame(post$var_anc)
names(var_anc.post)<-"post.var_anc.sims"


var_anc.plot<-ggplot()+
  geom_density(aes(prior.var_anc.sims,fill="prior.var_anc.sims"),alpha=0.2,data=var_anc.sims)+
  geom_density(aes(post.var_anc.sims,fill="post.var_anc.sims"),alpha=0.2,data=var_anc.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="var_anc", y = "Density")+
  geom_vline(xintercept=c(var_anc),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

var_anc.plot
```

Sigma plot
```{r}
sigma.sims<-rexp(n=1000,rate=1)
sigma.sims<-data.frame(sigma.sims)
names(sigma.sims)<-"prior.sigma.sims"


sigma.post<-data.frame(post$sigma)
names(sigma.post)<-"post.sigma.sims"


sigma.plot<-ggplot()+
  geom_density(aes(prior.sigma.sims,fill="prior.sigma.sims"),alpha=0.2,data=sigma.sims)+
  geom_density(aes(post.sigma.sims,fill="post.sigma.sims"),alpha=0.2,data=sigma.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="Sigma", y = "Density")+
  #geom_vline(xintercept=c(var_anc),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

sigma.plot
```



Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;

#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.05) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:100){
    curve(var_anc.sims[i,] * exp(-a.sims[i,] * x) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:100){
    curve(post$var_anc[i]  * exp(-post$a[i] * x) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```



Prior vs. Posterior Plot - centered model
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)


optima_bar<-rnorm(n=100,0,1)
#sigma<-abs(rnorm(n=100,0,1))

optima.sims<-rnorm(optima_bar,sigma)
optima.sims<-data.frame(optima.sims,Prior="Prior")

optima.post<-data.frame(post$optima)
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,optima)
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=optima.sims,aes(y=optima.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),linewidth=25,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
#            theme(
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank())+

  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```


Prior vs. Posterior Plot - non-centered model
```{r}
library(ggsci)
library(rethinking)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)


optima_bar<-rlnorm(n=100,0,0.5)
z<-rnorm(n=100,0,1)
sigma<-abs(rnorm(n=100,0,1))
#sigma<-rexp(n=100,rate=5)

optima.sims<-optima_bar + z*sigma;
optima.sims<-data.frame(optima.sims,Prior="Prior")

optima.post<-data.frame(post$optima)
names(optima.post)<-c(paste("OU",1:dim(optima.post)[2],sep=""))

#mu.link<-function(x.seq){alpha.post+x.seq*beta.post}
#x.seq <- seq(from=min(X), to=max(X) , length.out=100)
#mu <- sapply(x.seq , mu.link )
mu.mean <-apply( optima.post, 2, mean )
mu.mean<-data.frame(mu.mean)
mu.mean<-data.frame(mu.mean,"Regimes"=1:dim(mu.mean)[1])
mu.mean<-data.frame(mu.mean,optima)
#mu.mean<-data.frame(as.numeric(mu.mean))
#mu.mean<-data.frame(t(mu.mean),names(mu.mean))
#names(mu.mean)[3]<-"Regime"

mu.CI <- apply( optima.post , 2, PI , prob=0.89 )
mu.CI<-data.frame(t(data.frame(mu.CI)),"Regimes"=1:dim(mu.CI)[2])
names(mu.CI)<-c("min.5.5","max.94.5","Regimes")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,regimes=regimes_tip)
names(df)<-c("Y","Regimes")



optima.plot<-
   ggplot(data=df, Mapping = aes(x = Y, y = Regimes))+

  geom_jitter(data=optima.sims,aes(y=optima.sims,x=Prior),alpha=0.25,width=0.15)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=mu.mean,yend=mu.mean),linetype=1)+
  geom_segment(data=mu.mean,aes(x=Regimes-0.5,xend=Regimes+0.5,y=optima,yend=optima),linetype=2)+
  geom_linerange(data=mu.CI, mapping=aes(x=Regimes,ymin=min.5.5,ymax=max.94.5),linewidth=25,color="darkgrey",alpha=0.5)+
  geom_jitter(data=df,aes(y=Y,x=Regimes),width=0.15)+

  theme_bw()+
#            theme(
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank())+

  ylab("Optima") + xlab("")+
  scale_color_npg()

optima.plot
```

Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, covariance.plot, optima.plot, ncol=2,nrow=2, labels = c("A)","B)","C)"),common.legend = TRUE,legend="top")
#ggsave("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS17.pdf", plot = fig, width=7, height=7 )
ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures/blouch_validation_figS47.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS47C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()


```


#############################################################################################################################
Milestone 9: Combination multilevel Regime Model with centered priors and direct effect model
Four regimes
Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"

hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```



Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.4) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```


Slope plots for direct model with measurement error
```{r}
library(ggsci)
library(rethinking)
X<-X_with_error
Y<-Y_with_error
sigma<-rexp(100,5);

optima_bar.sims<-rnorm(100,mean(Y_with_error),0.25)
optima.sims<-rnorm(100,optima_bar.sims,sigma)

beta.sims<-rnorm(n=100,0,0.25)

optima.post<-post$optima

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

mu.link.11<-function(x.seq){optima.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){optima.post[,2]+x.seq*beta.post.1}

mu.link.21<-function(x.seq){optima.post[,3]+x.seq*beta.post.1}
mu.link.22<-function(x.seq){optima.post[,4]+x.seq*beta.post.1}

x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.21 <- sapply(x.seq , mu.link.21 )
mu.22 <- sapply(x.seq , mu.link.22 )


mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )
mu.mean.21 <-lapply( mu.21 , mean )
mu.mean.22 <-lapply( mu.22 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"
mu.mean.21<-data.frame(as.numeric(mu.mean.21))
mu.mean.22<-data.frame(as.numeric(mu.mean.22))
names(mu.mean.21)<-"mu.mean.21"
names(mu.mean.22)<-"mu.mean.22"



mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)

mu.CI.21 <- lapply( mu.21 , PI , prob=0.89 )
mu.CI.22 <- lapply( mu.22 , PI , prob=0.89 )
mu.CI.21<-data.frame(t(data.frame(mu.CI.21)),x.seq)
mu.CI.22<-data.frame(t(data.frame(mu.CI.22)),x.seq)


names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.21)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.22)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)
df21<-data.frame(x.seq,mu.mean.21)
df22<-data.frame(x.seq,mu.mean.22)


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=optima.sims,slope=beta.sims,alpha=0.1)+ #Prior
  
  geom_abline(intercept=optima[1],slope=optima[5],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[2],slope=optima[5],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[3],slope=optima[5],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[4],slope=optima[5],alpha=0.5,linetype=2)+
  
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  
  geom_line(data=df21,aes(x=x.seq,y=mu.mean.21),linetype=1)+
  geom_ribbon(data=mu.CI.21,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df22,aes(x=x.seq,y=mu.mean.22),linetype=1)+
  geom_ribbon(data=mu.CI.22,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+

  
    theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1


```



Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, "",slope.plot.1, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS50.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Documents/Validation Figures//blouch_validation_figS50C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
```


#############################################################################################################################
Milestone 10: Combination Regime Model with direct effect model and varying slopes
Two regimes - blouchOU_reg_direct_ME_Vs.stan
Hl Plot prior vs. posterior - assume posterior has been extracted using extract(model) and stored in post
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

hl.post<-data.frame(post$hl)
names(hl.post)<-"post.hl.sims"

hl.plot<-ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  geom_density(aes(post.hl.sims,fill="post.hl.sims"),alpha=0.2,data=hl.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="Half-life", y = "Density")+
  labs(title="",x="Half-life", y = "Density")+

    #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(hl),linetype=2)+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

hl.plot
```



```{r}
vy.sims<-rexp(n=1000,rate=20)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"


vy.post<-data.frame(post$vy)
names(vy.post)<-"post.vy.sims"


vy.plot<-ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  geom_density(aes(post.vy.sims,fill="post.vy.sims"),alpha=0.2,data=vy.post)+
  theme_bw()+
            theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #labs(title="Prior vs. Posterior Distribution ",x="vy", y = "Density")+
  labs(title="",x="vy", y = "Density")+
  geom_vline(xintercept=c(vy),linetype=2)+

  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

vy.plot
```



Direct effect model
Prior on how shared branch lengths correspond to covariance
```{r}
a.sims<-log(2)/hl.sims;
sigma2_y.sims<-vy.sims*(2*(log(2)/hl.sims));
#x<-seq(from=0,to=1,by=0.001)
#V.sim<-calc_direct_V(phy,a.sims,sigma2_y.sims)


plot( NULL , xlim=c(0,1) , ylim=c(0,0.4) , xlab="Time since MRCA" , ylab="Covariance" ,cex.axis=0.75, mgp=c(1.25,0.25,0),tcl=-0.25)
for (i in 1:30){
    curve(sigma2_y.sims[i,] /(2 * a.sims[i,]) * ((1 - exp(-2 * a.sims[i,] * (1-(x/2)))) * exp(-a.sims[i,] * x)) , add=TRUE , lwd=4 , col=col.alpha(1,0.15))
}

for (i in 1:30){
    curve(post$sigma2_y[i] /(2 * post$a[i]) * ((1 - exp(-2 * post$a[i] * (1-(x/2)))) * exp(-post$a[i] * x)) , add=TRUE , lwd=4 , col=col.alpha(2,0.5))
}

par(mar=c(3,3,0.25,0.25))
covariance.plot <- recordPlot()
dev.off()

#Vt = sigma2_y /( 2 * a) * ((1 - exp(-2 * a * ta)) .* exp(-a * tij)); //From Hansen (1997)

covariance.plot
```



Two regimes - Slope plots for direct model with measurement error
```{r}
library(ggsci)
library(rethinking)
X<-X_with_error
Y<-Y_with_error

optima.sims<-rnorm(100,mean(Y),1)
beta.sims<-rnorm(n=100,0,0.25)

optima.post<-post$optima_beta
beta.post<-data.frame(post$beta)
names(beta.post)<-c("post.beta.1","post.beta.2")

mu.link.11<-function(x.seq){optima.post[,1]+x.seq*beta.post[,1]}
mu.link.12<-function(x.seq){optima.post[,2]+x.seq*beta.post[,2]}

x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )

#mu.mean.11 <-lapply( mu.11 , mean )
#mu.mean.12 <-lapply( mu.12 , mean )

mu.mean.11<-colMeans(mu.11)
mu.mean.12<-colMeans(mu.12)

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"

mu.CI.11 <- apply( mu.11 , MARGIN=2, FUN=PI , prob=0.89 )
mu.CI.12 <- apply( mu.12 , MARGIN=2, FUN=PI , prob=0.89 )


#mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
#mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)

names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=optima.sims,slope=beta.sims,alpha=0.1)+ #Prior
  
  geom_abline(intercept=optima[1],slope=beta[1],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[2],slope=beta[2],alpha=0.5,linetype=2)+

  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  
    theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1


```




Four regimes - Slope plots for direct model with measurement error
```{r}
library(ggsci)
library(rethinking)
X<-X_with_error
Y<-Y_with_error

optima.sims<-rnorm(100,mean(Y),1)
beta.sims<-rnorm(n=100,0,0.25)

optima.post<-post$optima.beta

beta.post.1<-data.frame(post$beta[,1])
names(beta.post.1)<-"post.beta.1"

mu.link.11<-function(x.seq){optima.post[,1]+x.seq*beta.post.1}
mu.link.12<-function(x.seq){optima.post[,2]+x.seq*beta.post.1}

mu.link.21<-function(x.seq){optima.post[,3]+x.seq*beta.post.1}
mu.link.22<-function(x.seq){optima.post[,4]+x.seq*beta.post.1}

x.seq <- seq(from=min(X), to=max(X) , length.out=100)
mu.11 <- sapply(x.seq , mu.link.11 )
mu.12 <- sapply(x.seq , mu.link.12 )
mu.21 <- sapply(x.seq , mu.link.21 )
mu.22 <- sapply(x.seq , mu.link.22 )


mu.mean.11 <-lapply( mu.11 , mean )
mu.mean.12 <-lapply( mu.12 , mean )
mu.mean.21 <-lapply( mu.21 , mean )
mu.mean.22 <-lapply( mu.22 , mean )

mu.mean.11<-data.frame(as.numeric(mu.mean.11))
mu.mean.12<-data.frame(as.numeric(mu.mean.12))
names(mu.mean.11)<-"mu.mean.11"
names(mu.mean.12)<-"mu.mean.12"
mu.mean.21<-data.frame(as.numeric(mu.mean.21))
mu.mean.22<-data.frame(as.numeric(mu.mean.22))
names(mu.mean.21)<-"mu.mean.21"
names(mu.mean.22)<-"mu.mean.22"



mu.CI.11 <- lapply( mu.11 , PI , prob=0.89 )
mu.CI.12 <- lapply( mu.12 , PI , prob=0.89 )
mu.CI.11<-data.frame(t(data.frame(mu.CI.11)),x.seq)
mu.CI.12<-data.frame(t(data.frame(mu.CI.12)),x.seq)

mu.CI.21 <- lapply( mu.21 , PI , prob=0.89 )
mu.CI.22 <- lapply( mu.22 , PI , prob=0.89 )
mu.CI.21<-data.frame(t(data.frame(mu.CI.21)),x.seq)
mu.CI.22<-data.frame(t(data.frame(mu.CI.22)),x.seq)


names(mu.CI.11)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.12)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.21)<-c("min.5.5","max.94.5","x.seq")
names(mu.CI.22)<-c("min.5.5","max.94.5","x.seq")

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=dat$Y_obs,X=dat$X_obs,Regimes=regimes_tip)
df11<-data.frame(x.seq,mu.mean.11)
df12<-data.frame(x.seq,mu.mean.12)
df21<-data.frame(x.seq,mu.mean.21)
df22<-data.frame(x.seq,mu.mean.22)


#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.plot.1<-ggplot()+  
  geom_point(data=df,aes(y=Y,x=X,color=Regimes))+
  geom_abline(intercept=optima.sims,slope=beta.sims,alpha=0.1)+ #Prior
  
  geom_abline(intercept=optima[1],slope=optima[5],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[2],slope=optima[5],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[3],slope=optima[5],alpha=0.5,linetype=2)+
  geom_abline(intercept=optima[4],slope=optima[5],alpha=0.5,linetype=2)+
  
  geom_line(data=df11,aes(x=x.seq,y=mu.mean.11),linetype=1)+
  geom_ribbon(data=mu.CI.11,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df12,aes(x=x.seq,y=mu.mean.12),linetype=1)+
  geom_ribbon(data=mu.CI.12,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  
  geom_line(data=df21,aes(x=x.seq,y=mu.mean.21),linetype=1)+
  geom_ribbon(data=mu.CI.21,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+
  geom_line(data=df22,aes(x=x.seq,y=mu.mean.22),linetype=1)+
  geom_ribbon(data=mu.CI.22,aes(x=x.seq,ymin=min.5.5,ymax=max.94.5),linetype=2,alpha=0.25)+

  
    theme_bw()+
        theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())+

  #ggtitle("Prior vs. Posterior for Intercept and Slope")+
  ylab("Y") + xlab("Direct effect trait")+
  scale_color_npg()

slope.plot.1


```



Change figure number depending on model/parameter combination
```{r}
fig<-ggarrange(hl.plot, vy.plot, "",slope.plot.1, ncol=2,nrow=2, labels = c("A)","B)","C)","D)"),common.legend = TRUE,legend="top")

ggsave("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/blouch - not online/Validation Figures/blouch_validation_figS53.pdf", plot = fig, width=7, height=7 )

pdf(file = "/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/blouch - not online/Validation Figures/blouch_validation_figS53C.pdf",   # The directory you want to save the file in
    width = 3.6, # The width of the plot in inches
    height = 3.4) # The height of the plot in inches
covariance.plot
dev.off()
```



















##################################################################################################






Half life for priors - asssume phylogeny is scaled to unit length
```{r}
hl.sims<-rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75)
quantile(hl.sims,probs=c(0.1,0.25,0.5,0.75,0.9))

hl.sims[hl.sims<0.1]
hl.sims[hl.sims>1]

```

Prior Exploration
```{r}
hl.sims<-data.frame(rlnorm(n=1000,meanlog=log(0.25),sdlog=0.75))
#hl.sims<-data.frame(hl.sims[hl.sims<3])
names(hl.sims)<-"prior.hl.sims"

ggplot()+
  geom_density(aes(prior.hl.sims,fill="prior.hl.sims"),alpha=0.2,data=hl.sims)+
  theme_bw()+
  labs(title="Prior Distribution ",x="Half-life", y = "Density")+
  #scale_fill_manual(labels=c("Posterior","Prior"))+
  geom_vline(xintercept=c(0.1,1),linetype=2)
  scale_fill_npg(name="",labels=c("Prior"))

  quantile(hl.sims$prior.hl.sims,probs=c(0.1,0.25,0.5,0.75,0.9))
  
  

```


Sigma2y - asssume phylogeny is scaled to unit length
```{r}
sigma2_y<-rexp(n=1000,rate=5)
plot(density(sigma2_y),main="Sigma2_y prior distribution")
quantile(sigma2_y,probs=c(0.1,0.5,0.75,0.9))

sigma2_y[sigma2_y<0.1]
sigma2_y[sigma2_y>1]
```



Plot prior vs. posterior
```{r}
sigma2y.sims<-rexp(n=1000,rate=5)
sigma2y.sims<-data.frame(sigma2y.sims)
names(sigma2y.sims)<-"prior.sigma2y.sims"


sigma2y.post<-data.frame(post$sigma2_y)
names(sigma2y.post)<-"post.sigma2y.sims"


ggplot()+
  geom_density(aes(prior.sigma2y.sims,fill="prior.sigma2y.sims"),alpha=0.2,data=sigma2y.sims)+
  geom_density(aes(post.sigma2y.sims,fill="post.sigma2y.sims"),alpha=0.2,data=sigma2y.post)+
  theme_bw()+
  labs(title="",x="Sigma2_y", y = "Density")+
  #labs(title="Prior vs. Posterior Distribution ",x="Sigma2_y", y = "Density")+
  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior","Prior"))

```

vy
```{r}
vy.sims<-rexp(n=1000,rate=5)
vy.sims<-data.frame(vy.sims)
names(vy.sims)<-"prior.vy.sims"
quantile(vy.sims$prior.vy.sims,probs=c(0.1,0.5,0.75,0.9))


ggplot()+
  geom_density(aes(prior.vy.sims,fill="prior.vy.sims"),alpha=0.2,data=vy.sims)+
  theme_bw()+
  labs(title="Prior Distribution ",x="vy", y = "Density")+
  #scale_fill_manual(labels=c("Posterior","Prior"))+
    geom_vline(xintercept=c(vy),linetype=2)+

  scale_fill_npg(name="",labels=c("Prior"))

```




Optima from Regime-Only model
```{r}
optima.sims<-rnorm(n=1000,0,1)
optima.sims<-data.frame(optima.sims)
names(optima.sims)<-"prior.optima.sims"

optima.post<-data.frame(post$optima)
names(optima.post)<-c("post.optima.1","post.optima.2")


ggplot()+
  geom_density(aes(prior.optima.sims,fill="prior.optima.sims"),alpha=0.2,data=optima.sims)+
  geom_density(aes(post.optima.1,fill="post.optima.1"),alpha=0.2,data=optima.post)+
  geom_density(aes(post.optima.2,fill="post.optima.2"),alpha=0.2,data=optima.post)+

    theme_bw()+
  labs(title="Prior vs. Posterior Distribution ",x="Optima", y = "Density")+
  #scale_fill_manual(labels=c("Posterior","Prior"))+
  scale_fill_npg(name="",labels=c("Posterior Optima 1","Posterior Optima 2","Prior"))

```











Slope and intercept - simple one level no regimes model
```{r}
library(ggsci)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

#Adaptive Model
intercept_test<-rnorm(100,4,0.2)
#slope_test<-rnorm(100,0,0.2)
slope_test<-rlnorm(n=100,meanlog=log(0.4),sdlog=1)

#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=Y,X=X)

names(df)<-c("Y","X")

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.prior.plot<-ggplot(data=df,aes(y=Y,x=X))+  
  geom_point()+
  geom_abline(intercept=intercept_test,slope=slope_test,alpha=0.25)+
  theme_bw()+
  ggtitle("Priors for Intercept and Slope")+
  ylab("Trait 1") + xlab("Trait 2")+
  scale_color_npg()

slope.prior.plot
```

slope and intercept - Two Xs
```{r}
library(ggsci)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

#Adaptive Model
intercept_test<-rnorm(100,4,0.2)
slope_test<-rnorm(100,0,0.2)

df<-data.frame(Y=Y,X=Xs)

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.prior.plot<-ggplot(data=df,aes(y=Y,x=X.1))+  
  geom_point()+
  geom_abline(intercept=intercept_test,slope=slope_test,alpha=0.25)+
  theme_bw()+
  ggtitle("Priors for Intercept and Slope")+
  ylab("Trait 1") + xlab("Trait 2")+
  scale_color_npg()

slope.prior.plot
#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.prior.plot<-ggplot(data=df,aes(y=Y,x=X.2))+  
  geom_point()+
  geom_abline(intercept=intercept_test,slope=slope_test,alpha=0.25)+
  theme_bw()+
  ggtitle("Priors for Intercept and Slope")+
  ylab("Trait 1") + xlab("Trait 3")+
  scale_color_npg()

slope.prior.plot

```


Plot priors vs posterior
```{r}

```


Slope and intercept - simple one level regime model
```{r}
library(ggsci)
#Direct Model
#intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.2)
#slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)

#Adaptive Model
intercept_test<-rnorm(100,stan_sim_data$ols_intercept,0.1)
slope_test<-rnorm(100,stan_sim_data$ols_slope,0.1)


#df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$random_cov)

names(df)<-c("Y","X")

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.prior.plot<-ggplot(data=df,aes(y=Y,x=X))+  
  geom_point()+
  geom_abline(intercept=intercept_test,slope=slope_test,alpha=0.25)+
  theme_bw()+
  ggtitle("Priors for Intercept and Slope")+
  ylab("Trait 1") + xlab("Trait 2")+
  scale_color_npg()

slope.prior.plot
```


Multilevel model - Normal Y data
```{r}
#beta_raw<-rnorm(100,0,1)
beta_bar<-rnorm(100,stan_sim_data$ols_intercept,0.05)
sd_beta<-rexp(100,7)
beta<-rnorm(100,beta_bar,sd_beta)
slope_test<-rnorm(100,stan_sim_data$ols_slope,0.05)

df<-data.frame(Y=stan_sim_data$Y,X=stan_sim_data$direct_cov)
names(df)<-c("Y","X")

#slope.prior.plot<-ggplot(data=reg.trdata$dat,aes(y=Sim1,x=X))+
slope.prior.plot<-ggplot(data=df,aes(y=Y,x=X))+  
  geom_point()+
  geom_abline(intercept=beta,slope=slope_test,alpha=0.25)+
  theme_bw()+
  ggtitle("Priors for Intercept and Slope")+
  ylab("Trait 1") + xlab("Trait 2")+
  scale_color_npg()

slope.prior.plot
```

Regime only model
Plotting intercepts
