---
title: "Blouch SBR1 - Cervidae Tree Fixed Regime Data Simulation"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
Code to simulate fixed regimes on Cervidae Tree


Setup Workspace and load packages
```{r}
rm(list=ls())
library(devtools)
#devtools::install_github("Mark-Grabowski/blouch")

#library(blouch)
#load_all()
# Load necessary packages
library(ape)
library(slouch)
library(rstan)
library(treeplyr)
library(ggplot2)
library(ggsci)
library(bridgesampling)
library(mvSLOUCH)
#For execution on a local, multicore CPU with excess RAM we recommend calling
options(mc.cores = parallel::detectCores())
#options(mc.cores = 8)
rstan_options(auto_write = TRUE)

dotR <- file.path(Sys.getenv("HOME"), ".R")
if (!file.exists(dotR)) dir.create(dotR)
M <- file.path(dotR, "Makevars")
if (!file.exists(M)) file.create(M)
arch <- ifelse(R.version$arch == "aarch64", "arm64", "x86_64")
cat(paste("\nCXX14FLAGS += -O3 -mtune=native -arch", arch, "-ftemplate-depth-256"),
    file = M, sep = "\n", append = FALSE)

```

Data Prep
```{r}
cervid.tree<-read.nexus("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Sharable Data/cervidae_renamed.tre") #Macbook Pro

#cervid.tree<-read.nexus("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Sharable Data/cervidae_renamed.tre") #Mac Studio

cervid.dataset<-read.csv("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Sharable Data/updated_mean_dat_no_OL_MG.csv") #Macbook Pro

#cervid.dataset<-read.csv("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Sharable Data/updated_mean_dat_no_OL_MG.csv") #Mac Studio

#Remove Muntiacus_atherodes, Elaphodus_cephalophus - rudamentary and female Rangifer. 
cervid.dataset<-filter(cervid.dataset,Genus_Species != "Sinomegaloceros_yabei" & ! Genus_Species =="Alces_alces_gigas"  & ! Genus_Species == "Muntiacus_truongsonensis" & ! Genus_Species ==  "Mazama_temama"& ! Genus_Species ==  "Muntiacus_feae" & ! Genus_Species ==  "Muntiacus_atherodes" & ! Genus_Species ==  "Elaphodus_cephalophus")


cervid.trdata <- make.treedata(cervid.tree, cervid.dataset,name_column="Genus_Species")

cervid.trdata<-filter(cervid.trdata,!(is.na(log_ant_vol)) & !(is.na(log_psl)))
cervid.trdata<-mutate(cervid.trdata,mc.log_psl=cervid.trdata$dat$log_psl-(mean(cervid.trdata$dat$log_psl)))

cervid.trdata<-mutate(cervid.trdata,me.log_ant_vol=cervid.trdata$dat$log_vol_var_est/cervid.trdata$dat$n)
cervid.trdata<-mutate(cervid.trdata,me.log_psl=cervid.trdata$dat$log_psl_var_est/cervid.trdata$dat$n)
cervid.trdata<-filter(cervid.trdata,!(is.na(me.log_ant_vol)) & !(is.na(me.log_psl)))


cervid.trdata

```

Rescale Tree to Height =1
```{r}
l.tree<-max(branching.times(cervid.trdata$phy))
cervid.trdata$phy$edge.length<-cervid.trdata$phy$edge.length/l.tree ## rescale tree to height 1
max(branching.times(cervid.trdata$phy))

```

Set regimes - manually - 2 regimes
Locate nodes
```{r}
plot(cervid.trdata$phy)
tiplabels(cex=0.5)
nodelabels(cex=0.5)
#Using node 71
```


Set regimes - manually - 2 regimes
Locate nodes
```{r}
plot(cervid.trdata$phy)
tiplabels(cex=0.5)
nodelabels(cex=0.5)
#Using node 71
```

Set regimes
```{r}
source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Functions/set.converge.regimes.R") #Macbook Pro

#source("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Functions/set.converge.regimes.R") #Mac Studio

shifts<-c(71) #Location of nodes with regime shifts
cervid.2reg.trdata<-set.converge.regimes(cervid.trdata,shifts)


```

Check if manual setting code worked
```{r}
shifts.total<-c(cervid.2reg.trdata$dat$regimes,cervid.2reg.trdata$phy$node.label)
edge.regimes <- factor(shifts.total[cervid.2reg.trdata$phy$edge[,2]])
print(edge.regimes)
#Get ggplot colors used for plot to make on tree
gg_color_hue <- function(n) {
  hues = seq(15, 375, length=n+1)
  hcl(h=hues, l=65, c=100)[1:n]
}

reg.colors<-gg_color_hue(length(unique(cervid.2reg.trdata$dat$regimes)))

print(reg.colors)
plot(cervid.2reg.trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)


```


```{r}
#Direct and Random model - 1 trait each
name.response.trait<-c("log_ant_vol")
names.adaptive.traits<-c("mc.log_psl")
names.direct.traits<-c("mc.log_psl")
name.response.me.trait<-c("me.log_ant_vol")
names.adaptive.me.traits<-c("me.log_psl")
names.direct.me.traits<-c("me.log_psl")

names.fixed.fact<-c("regimes")

```



Simulate X and Y variable for Single Covariates - fixed Regimes/direct effect model
```{r}
source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/R Scripts/blouch_simXYOUReg.setup.R") #Macbook Pro

#source("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/R Scripts/blouch_simXYOUReg.setup.R") #Mac Studio

Sxx<-1
vX0<-0
optima<-c(4,5,0.8)

#One direct covariate w/o ME
stan_sim_data<-blouchsimXYOUReg.setup(cervid.2reg.trdata,names.fixed.fact,num.direct=1,num.random=0,hl=0.1,vy=0.01,beta=optima)

cervid.2reg.trdata$dat<-cbind(cervid.2reg.trdata$dat,stan_sim_data$direct_cov)
```


Simulate data
```{r}
setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/") #Macbook Pro
stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/simRegimes_XY.stan") #Macbook Pro


#setwd("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/")#Mac Studio
#stanc("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/simRegimes_XY.stan") #Mac Studio

stan_model <- stan_model("simRegimes_XY.stan")

fit.simdata<- rstan::sampling(object = stan_model,data = stan_sim_data,chains = 2,iter = 2000,control=list(adapt_delta=0.95),algorithm=c("Fixed_param"))

#For downstream analysis and plots
ext.fit.simdata <- rstan::extract(fit.simdata)

```


Add simulated code to dataset
```{r}
#Extract simulated brain mass data
Y_pred_sim<-as.data.frame(ext.fit.simdata$Y_pred)

#names(brain_mass_g_log_mean.sim)<-paste(names(brain_mass_g_log_mean.sim)

test<-data.frame(t(Y_pred_sim[1:100,]))
names(test)<-paste("Sim",1:100,sep="")

#Add to original dataset
cervid.2reg.trdata$dat<-cbind(cervid.2reg.trdata$dat,test)
```

Plot simulated relationships
```{r}
ggplot(data=cervid.2reg.trdata$dat,aes(x=trait_1,y=Sim1))+
  geom_point()

test<-cervid.2reg.trdata$dat$Sim1
names(test)<-cervid.2reg.trdata$phy$tip.label
dotTree(cervid.2reg.trdata$phy,test,length=10,ftype="i",fsize=0.5)

```

Set regimes - manually - 3 regimes
Locate nodes
```{r}
plot(cervid.trdata$phy)
tiplabels(cex=0.5)
nodelabels(cex=0.5)
#Using node 71
```

Set regimes
```{r}
source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Functions/set.converge.regimes.R") #Macbook Pro

#source("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Functions/set.converge.regimes.R") #Mac Studio


shifts<-c(71,52) #Location of nodes with regime shifts
cervid.3reg.trdata<-set.converge.regimes(cervid.trdata,shifts)


```

Check if manual setting code worked
```{r}
shifts.total<-c(cervid.3reg.trdata$dat$regimes,cervid.3reg.trdata$phy$node.label)
edge.regimes <- factor(shifts.total[cervid.3reg.trdata$phy$edge[,2]])
print(edge.regimes)
#Get ggplot colors used for plot to make on tree
gg_color_hue <- function(n) {
  hues = seq(15, 375, length=n+1)
  hcl(h=hues, l=65, c=100)[1:n]
}

reg.colors<-gg_color_hue(length(unique(cervid.3reg.trdata$dat$regimes)))

print(reg.colors)
plot(cervid.3reg.trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)


```



Simulate X and Y variable for Single Covariates - fixed Regimes/direct effect model
```{r}
source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/R Scripts/blouch_simXYOUReg.setup.R") #Macbook Pro

#source("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/R Scripts/blouch_simXYOUReg.setup.R") #Mac Studio


Sxx<-1
vX0<-0
optima<-c(4,5,6,0.8)

#One direct covariate w/o ME
stan_sim_data<-blouchsimXYOUReg.setup(cervid.3reg.trdata,names.fixed.fact,num.direct=1,num.random=0,hl=0.1,vy=0.01,beta=optima)

cervid.3reg.trdata$dat<-cbind(cervid.3reg.trdata$dat,stan_sim_data$direct_cov)
```


Simulate data
```{r}
setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/")
stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/simRegimes_XY.stan")

#setwd("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/")#Mac Studio
#stanc("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/simRegimes_XY.stan") #Mac Studio


stan_model <- stan_model("simRegimes_XY.stan")

fit.simdata<- rstan::sampling(object = stan_model,data = stan_sim_data,chains = 2,iter = 2000,control=list(adapt_delta=0.95),algorithm=c("Fixed_param"))

#For downstream analysis and plots
ext.fit.simdata <- rstan::extract(fit.simdata)

```


Add simulated code to dataset
```{r}
#Extract simulated brain mass data
Y_pred_sim<-as.data.frame(ext.fit.simdata$Y_pred)

#names(brain_mass_g_log_mean.sim)<-paste(names(brain_mass_g_log_mean.sim)

test<-data.frame(t(Y_pred_sim[1:100,]))
names(test)<-paste("Sim",1:100,sep="")

#Add to original dataset
cervid.3reg.trdata$dat<-cbind(cervid.3reg.trdata$dat,test)
```

Plot simulated relationships
```{r}
ggplot(data=cervid.3reg.trdata$dat,aes(x=trait_1,y=Sim1))+
  geom_point()

test<-cervid.3reg.trdata$dat$Sim1
names(test)<-cervid.3reg.trdata$phy$tip.label
dotTree(cervid.3reg.trdata$phy,test,length=10,ftype="i",fsize=0.5)

```


Set regimes - manually - 4 regimes
Locate nodes
```{r}
plot(cervid.trdata$phy)
tiplabels(cex=0.5)
nodelabels(cex=0.5)
#Using node 71
```

Set regimes
```{r}
source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Functions/set.converge.regimes.R") #Macbook Pro

#source("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Functions/set.converge.regimes.R") #Mac Studio


shifts<-c(71,52,59) #Location of nodes with regime shifts
cervid.4reg.trdata<-set.converge.regimes(cervid.trdata,shifts)


```

Check if manual setting code worked
```{r}
shifts.total<-c(cervid.4reg.trdata$dat$regimes,cervid.4reg.trdata$phy$node.label)
edge.regimes <- factor(shifts.total[cervid.4reg.trdata$phy$edge[,2]])
print(edge.regimes)
#Get ggplot colors used for plot to make on tree
gg_color_hue <- function(n) {
  hues = seq(15, 375, length=n+1)
  hcl(h=hues, l=65, c=100)[1:n]
}

reg.colors<-gg_color_hue(length(unique(cervid.4reg.trdata$dat$regimes)))

print(reg.colors)
plot(cervid.4reg.trdata$phy,edge.color = reg.colors[edge.regimes], edge.width = 1, cex = 0.2)


```


Simulate X and Y variable for Single Covariates - fixed Regimes/direct effect model
```{r}
source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/R Scripts/blouch_simXYOUReg.setup.R") #Macbook Pro

#source("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/R Scripts/blouch_simXYOUReg.setup.R") #Mac Studio

Sxx<-1
vX0<-0
optima<-c(3,4,5,6,0.8)

#One direct covariate w/o ME
stan_sim_data<-blouchsimXYOUReg.setup(cervid.4reg.trdata,names.fixed.fact,num.direct=1,num.random=0,hl=0.1,vy=0.01,beta=optima)

cervid.4reg.trdata$dat<-cbind(cervid.4reg.trdata$dat,stan_sim_data$direct_cov)
```


Simulate data
```{r}
setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/") #Macbook Pro
stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/simRegimes_XY.stan") #Macbook Pro

#setwd("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/")#Mac Studio
#stanc("/Users/markgrabowski/Library/CloudStorage/GoogleDrive-mark.walter.grabowski@gmail.com/Other computers/My MacBook Pro/Documents/Academic/Research/Current Projects/Blouch project/R1 blouch-testing branch/Simulation Code/Stan Models/simRegimes_XY.stan") #Mac Studio

stan_model <- stan_model("simRegimes_XY.stan")

fit.simdata<- rstan::sampling(object = stan_model,data = stan_sim_data,chains = 2,iter = 2000,control=list(adapt_delta=0.95),algorithm=c("Fixed_param"))

#For downstream analysis and plots
ext.fit.simdata <- rstan::extract(fit.simdata)

```


Add simulated code to dataset
```{r}
#Extract simulated brain mass data
Y_pred_sim<-as.data.frame(ext.fit.simdata$Y_pred)

#names(brain_mass_g_log_mean.sim)<-paste(names(brain_mass_g_log_mean.sim)

test<-data.frame(t(Y_pred_sim[1:100,]))
names(test)<-paste("Sim",1:100,sep="")

#Add to original dataset
cervid.4reg.trdata$dat<-cbind(cervid.4reg.trdata$dat,test)
```

Plot simulated relationships
```{r}
ggplot(data=cervid.4reg.trdata$dat,aes(x=trait_1,y=Sim1))+
  geom_point()

test<-cervid.4reg.trdata$dat$Sim1
names(test)<-cervid.4reg.trdata$phy$tip.label
dotTree(cervid.4reg.trdata$phy,test,length=10,ftype="i",fsize=0.5)

```




