---
title: "Blouch ms Cervid Analysis - Example"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
rm(list=ls())
library(rstan)
library(treeplyr)
#library(slouch)
#For execution on a local, multicore CPU with excess RAM we recommend calling
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

```

Data Prep
```{r}
cervid.tree<-read.nexus("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Data/Antler Dataset/cervidae_renamed.tre")


cervid.dataset<-read.csv("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Data/Antler Dataset/deer_mean_dat.csv")


#Remove Muntiacus_atherodes, Elaphodus_cephalophus - rudamentary and female Rangifer. 
cervid.dataset<-filter(cervid.dataset,Genus_Species != "Sinomegaloceros_yabei" & ! Genus_Species =="Alces_alces_gigas"  & ! Genus_Species == "Muntiacus_truongsonensis" & ! Genus_Species ==  "Mazama_temama"& ! Genus_Species ==  "Muntiacus_feae" & ! Genus_Species ==  "Muntiacus_atherodes" & ! Genus_Species ==  "Elaphodus_cephalophus")


cervid.trdata <- make.treedata(cervid.tree, cervid.dataset,name_column="Genus_Species")
cervid.trdata<-mutate(cervid.trdata,me.log_ant_vol=cervid.trdata$dat$log_vol_var_est/cervid.trdata$dat$n)
cervid.trdata<-mutate(cervid.trdata,me.log_psl=cervid.trdata$dat$log_psl_var_est/cervid.trdata$dat$n)

names.traits<-c("log_ant_vol","log_psl",NA,NA)

names.traits<-c("log_ant_vol","log_psl","me.log_ant_vol","me.log_psl")
cervid.trdata<-filter(cervid.trdata,!(is.na(log_ant_vol)) & !(is.na(log_psl)))
cervid.trdata<-filter(cervid.trdata,!(is.na(me.log_ant_vol)) & !(is.na(me.log_psl)))
cervid.trdata

#cervid.trdata<-filter(cervid.trdata, (Status == "Extant")) #Only non-fossil species

#Mean Standardized
cervid.trdata$dat$log_psl<-cervid.trdata$dat$log_psl-mean(cervid.trdata$dat$log_psl)




```


Rescale Tree to Height =1
```{r}
l.tree<-max(branching.times(cervid.trdata$phy))
cervid.trdata$phy$edge.length<-cervid.trdata$phy$edge.length/l.tree ## rescale tree to height 1
max(branching.times(cervid.trdata$phy))

```

Slouch Results for Comparison
Adaptation model
```{r}
#setwd("/Users/markgrabowski/Documents/Academic/Research/Approaches/Slouch/slouch-master")
#devtools::load_all()
#library(slouch)

#OU Model of Evolution
slouch.OU<-slouch.fit(phy = cervid.trdata$phy,
                              species = cervid.trdata$phy$tip.label,
                              response = cervid.trdata$dat$log_ant_vol,
                              random.cov = cervid.trdata$dat$log_psl,
                              mv.response = cervid.trdata$dat$me.log_ant_vol,
                              mv.random.cov = cervid.trdata$dat$me.log_psl,
                              hl_values = seq(0.00001, 0.2, length.out = 50),
                              vy_values = seq(0.00001, 0.5, length.out = 50),
                              hillclimb = TRUE,convergence = 150,
                              lower = c(0.00001, 0.00001)

                       )
summary(slouch.OU)
```
Constraint Model
```{r}
#setwd("/Users/markgrabowski/Documents/Academic/Research/Approaches/Slouch/slouch-master")
#devtools::load_all()
#library(slouch)

#OU Model of Evolution
slouch.OU<-slouch.fit(phy = cervid.trdata$phy,
                              species = cervid.trdata$phy$tip.label,
                              response = cervid.trdata$dat$log_ant_vol,
                              direct.cov = cervid.trdata$dat$log_psl,
                              mv.response = cervid.trdata$dat$me.log_ant_vol,
                              mv.direct.cov = cervid.trdata$dat$me.log_psl,
                              hl_values = seq(0.00001, 6, length.out = 50),
                              vy_values = seq(0.00001, 4, length.out = 50),
                              hillclimb = TRUE,convergence = 150,
                              lower = c(0.00001, 0.00001)

                       )
summary(slouch.OU)
```


Get bias correct coefficients
```{r}
slouch.OU$beta_primary$coefficients_bias_corr

```



Running Blouch Data Setup Function - 1st trait is adapting to 2
```{r}
par(mfrow = c(1,1))


#source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1/Blouch Setup Files/blouchOU.setup.v1.R")

source("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1/Blouch Setup Files/blouchOU.setup.v1.R")

#Data must be logged before entry into blouch.setup
#names.traits = c(response, predictor, me.response, me.predictor)
stan_data<-blouchOU.setup.v1(cervid.trdata,names.traits)

stan_constraint_data<-stan_data[[1]]
stan_adaptive_data<-stan_data[[2]]

```


Explore Priors
Slope and intercept
```{r}
library(ggsci)

intercept_test<-rnorm(100,stan_adaptive_data$ols_intercept,0.5)
#slope_test<-rnorm(100,ols.slope,0.5)
slope_test<-rnorm(100,stan_adaptive_data$ols_slope,1.0)

slope.prior.plot<-ggplot(data=cervid.trdata$dat,aes(y=log_ant_vol,x=log_psl))+
  geom_point()+
  geom_abline(intercept=intercept_test,slope=slope_test,alpha=0.25)+
  theme_bw()+
  ggtitle("Priors for Intercept and Slope")+
  ylab("log Antler Volume (l)") + xlab("log Posterior Skull Length (mm)")+
  scale_color_npg()

slope.prior.plot

```

```{r}
pdf("slope.prior.plot.pdf", width=5, height=5)
slope.prior.plot
dev.off()  

```



```{r}
###########################################
#Normal - based on alpha
#par.alpha <- list(meanlog=1.25, sdlog=0.75)
#par.alpha <- list(meanlog=1.0, sdlog=1.0) #Former runs 08/10/2021
par.alpha <- list(meanlog=0.75, sdlog=0.75) #Former runs 08/10/2021

samp <- rlnorm(10000, mean=par.alpha$meanlog, sd=par.alpha$sdlog)
samp<-samp[samp>=0]

hist(samp, breaks=10000, main="Prior density of alpha",xlim=c(0,3))
#abline(v=(c(log(2)/0.07142857,log(2)/1)), col="red", lwd=2, lty=2) #Lines for 1 Ma and length of tree (14 Ma) - 
min(samp)
max(samp)
log(2)/min(samp)
log(2)/max(samp)

quantiles <- c(0, 0.01, 0.025,0.10, 0.25, 0.5, 0.75,0.90,0.95, 0.975, 0.99, 1)
#hls <- rlnorm(10000, meanlog=log(2)/par.alpha$meanlog, sdlog=log(2)/par.alpha$sdlog)
#hls<-hls[hls>=0]
qs <- quantile(log(2)/samp, quantiles) ## Use 'alfs' and math to calculate the quantiles of phylogenetic half-life
round(qs, 2)

hist((log(2)/samp), breaks=10000, main="Prior density of half-life",xlim=c(0,3))
abline(v=(c(0.07142857,1)), col="red", lwd=2, lty=2) #Lines for 1 Ma and length of tree (14 Ma) - 
########################################
```



```{r}
library(ggsci)
hl.prior.plot<-ggplot(data.frame(half.life = log(2)/samp),aes(x=half.life))+
  geom_histogram(binwidth=0.01)+
  geom_vline(xintercept=c(0.07142857,1),colour="#BB0000",linetype="dashed")+
  theme_bw()+
  ggtitle("Prior for Half-life")+
  ylab("Frequency") + xlab("Half-life")+xlim(c(0,3))+
  scale_color_npg()
 
hl.prior.plot

pdf("hl.prior.plot.pdf", width=5, height=5)
hl.prior.plot
dev.off()
```




Sigma2_y Prior
```{r}
library(extraDistr)
###########################################
#Normal - based on alpha
par.sigma2y <- list(sigma=0.1)

samp <- rhcauchy(10000, sigma=par.sigma2y$sigma)

hist(samp, breaks=10000, main="Prior density of Sigma2y",xlim=c(0,3))
#abline(v=(c(log(2)/0.07142857,log(2)/1)), col="red", lwd=2, lty=2) #Lines for 1 Ma and length of tree (14 Ma) - 
min(samp)
max(samp)
log(2)/min(samp)
log(2)/max(samp)

quantiles <- c(0, 0.01, 0.025,0.10, 0.25, 0.5, 0.75, 0.90, 0.975, 0.99, 1)
#hls <- rlnorm(10000, meanlog=log(2)/par.alpha$meanlog, sdlog=log(2)/par.alpha$sdlog)
#hls<-hls[hls>=0]
qs <- quantile(samp, quantiles) ## Use 'alfs' and math to calculate the quantiles of phylogenetic half-life
round(qs, 2)

hist(samp/(2*2), breaks=10000, main="Prior density of Vy",xlim=c(0,3))
#abline(v=(c(0.07142857,1)), col="red", lwd=2, lty=2) #Lines for 1 Ma and length of tree (14 Ma) - 
########################################

```


```{r}
library(extraDistr)
###########################################
#Normal - based on alpha
par.sigma2y <- list(sigma=1.0)

samp <- rhcauchy(10000, sigma=par.sigma2y$sigma)

hist(samp, breaks=10000, main="Prior density of Sigma2y",xlim=c(0,3))
#abline(v=(c(log(2)/0.07142857,log(2)/1)), col="red", lwd=2, lty=2) #Lines for 1 Ma and length of tree (14 Ma) - 
min(samp)
max(samp)
log(2)/min(samp)
log(2)/max(samp)

quantiles <- c(0, 0.01, 0.025,0.10, 0.25, 0.5, 0.75, 0.90, 0.975, 0.99, 1)
#hls <- rlnorm(10000, meanlog=log(2)/par.alpha$meanlog, sdlog=log(2)/par.alpha$sdlog)
#hls<-hls[hls>=0]
qs <- quantile(samp, quantiles) ## Use 'alfs' and math to calculate the quantiles of phylogenetic half-life
round(qs, 2)

hist(samp/(2*2), breaks=10000, main="Prior density of Vy",xlim=c(0,3))
#abline(v=(c(0.07142857,1)), col="red", lwd=2, lty=2) #Lines for 1 Ma and length of tree (14 Ma) - 
########################################

```

Constraint Model - with a ~ normal(1.0,1.0), sigma2y ~ uniform
```{r}
#setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1")
setwd("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1")
stanc("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1/blouchOU_test.stan")
stan_model <- stan_model("blouchOU_test.stan")

fit.uni.direct<- rstan::sampling(object = stan_model,data = stan_constraint_data,chains = 2,iter = 4000,control=list(adapt_delta=0.95))

  
print(fit.uni.direct,pars = c("a","hl","alpha","beta","vy","r_squared","sigma2_y"))

#For downstream analysis and plots
posterior.fit.uni.direct <- rstan::extract(fit.uni.direct)

```

Sigma uniform trace and density plots
```{r}
par(mfrow = c(1,3))

traceplot(fit.uni.direct,c("a","hl","alpha","beta","vy","sigma2_y"))
stan_dens(fit.uni.direct,c("a","hl","alpha","beta","vy","sigma2_y"))
#3X8
```


Constraint Model - with a ~ normal(1.25,1.0), sigma2y ~ hc(1.0)
```{r}
setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1")
stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1/blouchOU_test.stan")
stan_model <- stan_model("blouchOU_test.stan")

fit.hc.direct<- rstan::sampling(object = stan_model,data = stan_constraint_data,chains = 2,iter = 4000,control=list(adapt_delta=0.95))

print(fit.hc.direct,pars = c("a","hl","alpha","beta","vy","r_squared","sigma2_y"))


```

```{r}
par(mfrow = c(1,3))

traceplot(fit.hc.direct,c("a","hl","alpha","beta","vy","sigma2_y"))
stan_dens(fit.hc.direct,c("a","hl","alpha","beta","vy","sigma2_y"))
#3X8
```

Model Comparison using Bayes Factors
```{r}
library(bridgesampling)

lml.fit.uni.direct<-bridge_sampler(fit.uni.direct,silent=TRUE)
lml.fit.hc.direct<-bridge_sampler(fit.hc.direct,silent=TRUE)
#lml.fit.BM<-bridge_sampler(fit.BM,silent=TRUE)
#BF_att <- bridgesampling::bf(lml.fit.direct, lml.fit.BM)
BF_att <- bridgesampling::bf(lml.fit.direct, lml.fit.hc.direct)
BF_att

```
## Estimated Bayes factor in favor of lml_pupil over lml_pupil_null: 0.99924

We find that the data is 0.999 more likely under a model that assumes a positive effect of load than under a model that assumes no effect.


BM Model - with sigma2y ~ hc(1.0)
```{r}
setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1")
stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1/blouchBM_v1.stan")
stan_model <- stan_model("blouchBM_v1.stan")

fit.BM.hc<- rstan::sampling(object = stan_model,data = stan_constraint_data,chains = 2,iter = 4000,control=list(adapt_delta=0.95))

print(fit.BM.hc,pars = c("alpha","beta","r_squared","sigma2_y"))

```



BM Model - with sigma2y ~ uniform
```{r}
setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1")
stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1/blouchBM_v1.stan")
stan_model <- stan_model("blouchBM_v1.stan")

fit.BM.uni<- rstan::sampling(object = stan_model,data = stan_constraint_data,chains = 2,iter = 4000,control=list(adapt_delta=0.95))

print(fit.BM.uni,pars = c("alpha","beta","r_squared","sigma2_y"))

```


Model Comparison using Bayes Factors
```{r}
library(bridgesampling)
lml.fit.BM.hc<-bridge_sampler(fit.BM.hc,silent=TRUE)
lml.fit.BM.uni<-bridge_sampler(fit.BM.uni,silent=TRUE)
#lml.fit.hc.direct<-bridge_sampler(fit.hc.direct,silent=TRUE)
#lml.fit.BM<-bridge_sampler(fit.BM,silent=TRUE)
#BF_att <- bridgesampling::bf(lml.fit.direct, lml.fit.BM)
BF_att <- bridgesampling::bf(lml.fit.BM.uni, lml.fit.BM.hc)
BF_att

```

Direct vs. BM Model Comparison using BF
```{r}
library(bridgesampling)
lml.fit.uni.direct<-bridge_sampler(fit.uni.direct,silent=TRUE)
lml.fit.BM.uni<-bridge_sampler(fit.BM.uni,silent=TRUE)

BF_att <- bridgesampling::bf(lml.fit.uni.direct, lml.fit.BM.uni)
BF_att

```






Make plots for Figure
```{r}
library(ggsci)
library(ggplot2)

```

Fig S1
```{r}
old.par <- par(mar = c(0, 0, 0, 0))
par(old.par)


ggplot(cervid.trdata$dat,aes(y=log_ant_vol,x=log_psl))+
  geom_abline(intercept=mean(posterior.fit.uni.direct$alpha),slope = mean(posterior.fit.uni.direct$beta),lty=1)+
  geom_abline(intercept=-1.16,slope = 5.89,lty=2)+

  geom_point(aes(color=Tribe),size=2.0,alpha=0.8)+
  theme_bw()+
  theme(legend.position="bottom")+
  ylab("log Antler Volume (l)") + xlab("log Posterior Skull Length (mm)")+
  scale_color_npg()
 
#5.25 X 5

```





Rerun for predicting Irish Elk Antler length
Data Prep
```{r}
#cervid.tree<-read.nexus("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Data/Antler Dataset/cervidae_renamed.tre")
cervid.tree<-read.nexus("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Data/Antler Dataset/cervidae_renamed.tre")

#cervid.dataset<-read.csv("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Data/Antler Dataset/deer_mean_dat.csv")

cervid.dataset<-read.csv("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Data/Antler Dataset/deer_mean_dat.csv")



#Remove Muntiacus_atherodes, Elaphodus_cephalophus - rudamentary and female Rangifer. 
cervid.dataset<-filter(cervid.dataset,Genus_Species != "Sinomegaloceros_yabei" & ! Genus_Species =="Alces_alces_gigas"  & ! Genus_Species == "Muntiacus_truongsonensis" & ! Genus_Species ==  "Mazama_temama"& ! Genus_Species ==  "Muntiacus_feae" & ! Genus_Species ==  "Muntiacus_atherodes" & ! Genus_Species ==  "Elaphodus_cephalophus")


cervid.trdata <- make.treedata(cervid.tree, cervid.dataset,name_column="Genus_Species")
cervid.trdata<-mutate(cervid.trdata,me.log_ant_vol=cervid.trdata$dat$log_vol_var_est/cervid.trdata$dat$n)
cervid.trdata<-mutate(cervid.trdata,me.log_psl=cervid.trdata$dat$log_psl_var_est/cervid.trdata$dat$n)

#names.traits<-c("log_ant_vol","log_psl",NA,NA)

names.traits<-c("log_ant_vol","log_psl","me.log_ant_vol","me.log_psl")
cervid.trdata<-filter(cervid.trdata,!(is.na(log_ant_vol)) & !(is.na(log_psl)))
cervid.trdata<-filter(cervid.trdata,!(is.na(me.log_ant_vol)) & !(is.na(me.log_psl)))

#Only for reduced Cervini analysis
#cervid.trdata<-filter(cervid.trdata, (Tribe == "Cervini")) #Only non-fossil species

cervid.ext.trdata<-filter(cervid.trdata, (Status == "Extant")) #Only non-fossil species

#Mean Standardized
cervid.trdata$dat$log_psl<-cervid.trdata$dat$log_psl-mean(cervid.trdata$dat$log_psl)
cervid.ext.trdata$dat$log_psl<-cervid.ext.trdata$dat$log_psl-mean(cervid.ext.trdata$dat$log_psl)


```

Rescale Tree to Height =1
```{r}
l.tree<-max(branching.times(cervid.trdata$phy))
cervid.trdata$phy$edge.length<-cervid.trdata$phy$edge.length/l.tree ## rescale tree to height 1
max(branching.times(cervid.trdata$phy))

l.tree<-max(branching.times(cervid.ext.trdata$phy))
cervid.ext.trdata$phy$edge.length<-cervid.ext.trdata$phy$edge.length/l.tree ## rescale tree to height 1
max(branching.times(cervid.ext.trdata$phy))


```


Running Blouch Data Setup Function - 1st trait is adapting to 2
```{r}
#source("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1/Blouch Setup Files/blouchOUPredict.setup.v1.R")

source("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1/Blouch Setup Files/blouchOUPredict.setup.v1.R")

#Data must be logged before entry into blouch.setup
#names.traits = c(response, predictor, me.response, me.predictor) - classical = 0 for inverse regression
stan_data<-blouchOUPredict.setup.v1(cervid.ext.trdata,cervid.trdata,names.traits,classical=0)

stan_constraint_data<-stan_data[[1]]
stan_adaptive_data<-stan_data[[2]]

```

Constraint Model
```{r}
#setwd("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1")
#stanc("/Users/markgrabowski/Documents/Academic/Research/Current Projects/Blouch project/Stan Functions/blouch v1/blouchOUPredict_v1.stan")

setwd("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1")
#stanc("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1/blouchOUPredict_v1.stan")

stanc("/Users/markgrabowski/Google Drive/Shared with Macbook/Current Projects/Blouch project/Stan Functions/blouch v1/blouchOUPredict_test.stan")

#stan_model <- stan_model("blouchOUPredict_v1.stan")
stan_model <- stan_model("blouchOUPredict_test.stan")

#fit.direct<- stan(file = stan_model,data = stan_constraint_data,chains = 1,iter = 400,control=list(adapt_delta=0.95))
fit.fos.direct<- rstan::sampling(object = stan_model,data = stan_constraint_data,chains = 2,iter = 4000,control=list(adapt_delta=0.95),save_warmup=FALSE)

print(fit.fos.direct,pars = c("a","hl","alpha","beta","vy","r_squared","sigma2_y","RMSE","RMSE_mu","Y_pred_fos_means"))
posterior.fit.fos.direct <- rstan::extract(fit.fos.direct)

```


```{r}
par(mfrow = c(1,3))
  
traceplot(fit.fos.direct,c("a","hl","alpha","beta","vy","sigma2_y"))
stan_dens(fit.fos.direct,c("a","hl","alpha","beta","vy","sigma2_y"))
#3 X 8
```


Make plots for Figure
```{r}
library(ggsci)
library(ggplot2)
library(plotly)

```

Fig. 1
```{r}
cervid.fos.plot<-ggplot(cervid.trdata$dat,aes(y=log_ant_vol,x=log_psl))+
  geom_abline(intercept=mean(posterior.fit.fos.direct$alpha),slope = mean(posterior.fit.fos.direct$beta),lty=1)+
  #geom_abline(intercept=-15.123391,slope = 5.932,lty=2)+
  #geom_abline(intercept=-11.8719729343,slope = 4.86,lty=2)+ #Cervini slope going through Dama dama - from Tsuboi et al.
  #geom_abline(intercept=-11.48,slope = 4.70,lty=3)+ #Cervini slope going through Dama dama - from Blouch
  #geom_point(aes(y= 3.28,x=3.089545121),size=2.0,alpha=0.8)+
  geom_point(aes(y= 3.28,x=0.7373661),size=2.0,alpha=0.8)+
  geom_point(aes(color=Tribe),size=2.0,alpha=0.8)+
  theme_bw()+
  #theme(legend.position="bottom")+
  ylab("log Antler Volume (l)") + xlab("log Posterior Skull Length (mm)")+
  scale_color_npg()
 

cervid.fos.plot
#ggplotly()
#5.25 X 5
#5 X 7
```
Plot including predictions for extant species
Fig. 2b
SI HW Prediction Means - 5X5 pdf
```{r}
old.par <- par(mar = c(0, 0, 0, 0))
par(old.par)


fos.index<-which(cervid.trdata$dat$Status=="Extinct")
print(paste("Fossil Species #",fos.index))
extant.index<-which(cervid.trdata$dat$Status=="Extant")

antler.predictions<-apply(posterior.fit.fos.direct$Y_pred_fos_means,2,mean)
antler.predictions.extant<-apply(posterior.fit.fos.direct$Y_pred_extant_means,2,mean)

extant.data<-data.frame(Genus_Species = cervid.trdata$phy$tip.label[-fos.index],log_AV = cervid.trdata$dat$log_ant_vol[-fos.index],log_AV_predictions = antler.predictions.extant,log_psl = cervid.trdata$dat$log_psl[-fos.index],Status="Extant")

fos.predictions<-data.frame(Genus_Species = cervid.trdata$phy$tip.label[fos.index],log_AV = NA, log_AV_predictions = antler.predictions,log_psl=cervid.trdata$dat$log_psl[fos.index],Status="Extinct")
         
merged.data<-rbind(extant.data,fos.predictions)                   

antler.complete.plot<-ggplot(merged.data,aes(x=log_psl))+
geom_point(aes(y=log_AV),size=2.0,alpha=0.4)+
geom_point(aes(y=log_AV_predictions,color=Status),size=2.5,alpha=0.8)+
geom_abline(intercept=mean(posterior.fit.fos.direct$alpha),slope = mean(posterior.fit.fos.direct$beta),lty=2)+
theme_bw()+
theme(legend.position="bottom")+ #5X5
xlab("log Posterior Skull Length (mm)")+
  ylab("log Antler Volume (l)")

#+theme(legend. position = "none")
antler.complete.plot+scale_color_aaas()
#Export 5.25X5 PDF

```
